Using WordNet and Short-Term Memory for Contextual Disambiguation

### 2017-02-03 Draft_1.tex
Began work on disambiguation subsection of implementation section

        \Large
        Word Count - 3108\\
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

@@

If the activateSynset finds the synset in the STS,

 The most notable of its functions is sendToStm, shown in Listing \ref{lst:sendToStm}.

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={the sendToStm function}, captionpos=b, label={lst:sendToStm}]
FUNCTION sendToStm(inputSynset):
	episodicBuffer.remove(inputSynset)
	returnedItem := stm.swapLowestItem(inputSynset)
	episodicBuffer.add(returnedItem)
END FUNCTION
\end{lstlisting}

The swapLowestItem function, shown in Listing \ref{lst:swapLowestItem}, will return either the item sent (i.e. the STS rejects the item) or the item in the STS with the lowest activation. Whichever is returned will be added into the Episodic Buffer.

If the STS isn't full, the new item is simply added to the STS, and nothing is returned.

@@

\subsection{Forgetting}
\label{sec:forgetting}
After each sentence is read, all items inthe STS must be forgotten (have their activation reduced), so as to comply with Atkinson and Shiffrin's STS model discussed \hyperref[ShortTerm]{previously} \cite{ControlProcessesSTMAtkinson}. The forgetting of the entire contents of the STS is done using a loop, as shown in Listing \ref{lst:ForgetLoop}

@@

@@

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={Hypernym Activation}, captionpos=b, label={lst:HypernymActivation}]
FUNCTION activateHypernyms(synset, depth, activationModifier):
   activationModifier = hypernymModel(activationModifier)
   IF activationModifier < 0 THEN
       RETURN
   END IF
   activateSynset(synset, activationModifier)
   IF depth > 0 THEN
       FOR hypernym in synset.hypernyms() LOOP
           linearHypernym(hypernym, depth - 1, activationModifier)
       END LOOP
   ELSE:
       RETURN
   END IF
END FUNCTION
\end{lstlisting}

It can be seen that the function recursively activates the synsets of all hypernyms up to a specified depth. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

@@

\begin{figure}[h]
\begin{tikzpicture}
\begin{axis}[domain=0:10, grid=both, enlarge x limits=false]
\addplot[color=red]{log2(x)};
\end{axis}
\end{tikzpicture}
\caption{GRAPH OF $\log(x)$}
\label{fig:logGraph}
\end{figure}

One function which fits this criteria is $\log(x)$, shown in Figure \ref{fig:logGraph}. The gradient of the function, for higher values of $x$ is relatively shallow, favouring more general hypernyms heavily. For lower values of $x$, the gradient becomes very steep, causing more general hypernyms to have a lower activation increase.

It may be necessary to tune this function during experimentation, to provide greater accuracy, meaning more variables must be added. With these additions, the full model is:
\[f(x) = a + b\log(x)\]
where $a$ and $b$ are varied during experimentation, and $x$ is the

\subsection{Disambiguation}
\label{sec:ImplementedDisambiguation}

>>>>

        \Large
        Word Count - 4102\\
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

@@

If the activateSynset finds the synset in the STS, the synset is simply activated using, and remains there. If the synset is not present in the STS, it is activated, and sent to the STS, through the sendToStm function, shown in Listing \ref{lst:sendToStm}, recieving a boost if it has previously existed in the STS (i.e. it is in the Episodic Buffer).

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={the sendToStm function}, captionpos=b, label={lst:sendToStm}]
FUNCTION sendToStm(inputSynset):
	returnedItem = self.stm.swapLowestItem(inputItem)
    IF returnedItem is None THEN
        RETURN
    ELSE IF NOT episodicBuffer.inContents(returnedItem) THEN
        episodicBuffer.addSynset(returnedItem)
        RETURN
    END IF
END FUNCTION
\end{lstlisting}

The sendToStm function, shown in Listing \ref{lst:sendToStm}, is responsible for changing the contents of the STS and updating the Episodic buffer accordingly. The swapLowestItem function, shown in Listing \ref{lst:swapLowestItem} is used to edit the STS. As discussed \hyperref[sec:ImplementedSTS]{previously}, this function can return a synset, if the swap is successful. In this case, sendToStm will add the synset previously in the STS, to the episodic buffer (if it is not already present there).

@@

\subsection{Forgetting}
\label{sec:forgetting}
After each sentence is read, all items in the STS must be forgotten (have their activation reduced), so as to comply with Atkinson and Shiffrin's STS model discussed \hyperref[ShortTerm]{previously} \cite{ControlProcessesSTMAtkinson}. The forgetting of the entire contents of the STS is done using a loop, as shown in Listing \ref{lst:ForgetLoop}.

@@

The forgetting process occurs after every sentence, giving synsets in the next sentence the opportunity to enter the STS.

@@

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={Hypernym Activation}, captionpos=b, label={lst:HypernymActivation}]
FUNCTION activateHypernyms(synset, depth):
   activationModifier = hypernymModel(depth)
   IF activationModifier < 0 THEN
       RETURN
   ELSE
       FOR hypernym in synset.hypernyms() LOOP
           activateHypernyms(hypernym, depth + 1)
       END LOOP
       RETURN
   END IF
END FUNCTION
\end{lstlisting}

It can be seen that the function recursively activates the synsets of all hypernyms until, the model gives a result of less than zero. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

@@

\begin{figure}[h]
\begin{tikzpicture}
\begin{axis}[domain=0:10, grid=both, enlarge x limits=false]
\addplot[color=red]{-log2(x)};
\end{axis}
\end{tikzpicture}
\caption{GRAPH OF $-\log(x)$}
\label{fig:logGraph}
\end{figure}

One function which fits this criteria is $-\log(x)$, shown in Figure \ref{fig:logGraph}. The values given by the function, for higher values of $x$ are relatively high, favouring more general hypernyms heavily. For lower values of $x$, the values given are relatively low, causing more general hypernyms to have a lower activation increase.

It may be necessary to tune this function during experimentation, to provide greater accuracy, meaning more variables must be added. With these additions, the full model is:
\[f(x) = -b\log(ax)\]
where $a$ and $b$ are varied during experimentation, and $x$ is the depth of the hypernym. When the depth = 0 (i.e. the activation of the root synset) the model would require the calculation of log(0), which is undefined. For this reason, an offset of 0.001 was added, making the final model:
\[f(x) = -b\log(ax+0.001)\]
The size of the offset needed to be small to reduce its impact on the overall model.

\subsection{Disambiguation}
\label{sec:ImplementedDisambiguation}
Up to this point, we have only discussed the processes involved with changing the contents of the memory structures. The purpose of these memory structures is to provide the surrounding context of a word, to aid in its disambiguation.

As discussed in the \hyperref[sec:DisambiguationModels]{Disambiguation Models section}, the multiple access model relies upon context and frequency, using whichever is stronger for disambiguation \cite{PsychologyOfLanguage}. The system discussed in this section makes use of both of these, through the disambiguate function, shown in Listing \ref{lst:disambiguate}.

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={The disambiguate function}, captionpos=b, label={lst:disambiguate}]
FUNCTION disambiguate(synsetList):
    FOR item in stm LOOP
        IF item in synsetList THEN
            RETURN item
        END IF
    END LOOP
    FOR item in stm LOOP
        returnedSynset = hyponymSearch(synsetList, item)
        IF returnedSynset is not None THEN
            RETURN returnedSynset
        END IF
    END LOOP
    RETURN mostLikelySynset(synsetList)
END FUNCTION
\end{lstlisting}

\subsubsection{Context}
\label{sec:DisambiguationContext}
The context required for disambiguation exists in the STS. Initially, the disambiguate function traverses through the STS, searching for a synset of the word to be disambiguated. If successful, the function will return the found synset. If unsuccessful, more investigation is required, through the use of hyponyms.

As discussed in the \hyperref[sec:MattBurke]{Previous Works section}, M. Burke found that hyponym based searching, using the Stm is the most effective method of disambiguation \cite{MattBurkePrevious}. This method works upon the assumption that the contents of the Stm are relatively general (i.e. exist high up in the Wordnet hierarchy). The implementation of this search is given by the hyponymSearch function, shown in Listing \ref{lst:hyponymSearch}.

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={The hyponymSearch function}, captionpos=b, label={lst:hyponymSearch}]
FUNCTION hyponymSearch(synsetList, searchItem):
    hyponymList = searchItem.hyponyms()
    IF len(hyponymList) == 0 THEN
        RETURN None
    END IF
    FOR item in hyponymList LOOP
        IF item in synsetList THEN
            RETURN item
        END IF
    END LOOP
    FOR item in hyponymList LOOP
        returnedItem = hyponymSearch(synsetList, item)
        IF returnedItem is not None THEN
            RETURN returnedItem
        END IF
    END LOOP
    RETURN None
\end{lstlisting}

\subsubsection{Frequency}
\label{sec:DisambiguationFrequency}
